---
title: "Compile Emotrics with Demographics"
author: "Joelle Jee"
date: "January 28, 2022"
output:
  html_document:
    df_print: paged
---


```{r, message=FALSE}
if (!requireNamespace("tidyverse")) {
  install.packages("tidyverse")
}

if (!requireNamespace("kableExtra")) {
  install.packages("kableExtra")
}

if (!requireNamespace("knitr")) {
  install.packages("knitr")
}

library(tidyverse)
library(kableExtra)
library(knitr)

```

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE)
```

```{r}
project = file.path('/Users', Sys.info()['user'], 'Library', 
                    'CloudStorage', 'Box-Box', 
                    'Craniofacial Study', 'final_emotrics_f2g_analysis')

# read the most recent demographics master file
find_latest <- function(dir) {
  files = list.files(dir)
  files = gsub("\\D", "", files) %>% 
    as.Date('%m%d%y')
  
  latest = sort(files, decreasing = TRUE)[1]
  latest = format(latest, '%m%d%y')
  
}

latest = find_latest(file.path(project, 'demographics'))
master_demo <- list.files(file.path(project, 'demographics'))
master_demo = master_demo[endsWith(master_demo, paste0(latest, '.csv'))]
master_demo <- read.csv(file.path(project, 
                                  'demographics', 
                                  master_demo), 
                        check.names = FALSE)
master_demo$`In-lab Picture` = as.numeric(master_demo$`In-lab Picture`)
master_demo
```


# Merge Penn data with demogrpahics

## Load and rename Penn emotrics data columns

```{r}
latest = find_latest(file.path(project, 'Emotrics', 'tables'))
penn_emo = paste0('penn_final_emotrics_', latest, '.csv')
penn_emo <- read.csv(file.path(project, 
                               'Emotrics',
                               'tables',
                               latest,
                               penn_emo),
                     check.names = FALSE)


cols = sapply(1:ncol(penn_emo), 
             FUN=function(x) {
               col = colnames(penn_emo)[x]
               if (col == 'bbl_id'){
                 return(col)
               } else if (col == 'philtrum') {
                 return('Philtrum')
               } else {
                 col = paste(col, penn_emo[1, x], sep = '_')
                 col = str_replace_all(col, ' ', '_')
                 return(col)
               }
             })

colnames(penn_emo) = cols

penn_emo = penn_emo[,grepl('bbl_id|group|Right|Left|Philtrum', 
                           colnames(penn_emo))] %>%
  drop_na()

penn_emo

```

## Merge with master demograhpics file

### Penn subjects analyzed with Emotrics but missing demographics data

```{r}
penn_demo = master_demo[master_demo$group %in% c('CR', 'SZ', 'NC'),]
penn_emo$bbl_id[!penn_emo$bbl_id %in% penn_demo$bbl_id]
```


### Penn subjects with demographics data but missing Emotrics file

```{r}
penn_demo[!penn_demo$bbl_id %in% penn_emo$bbl_id, ]
```


```{r}
penn_merged = merge(master_demo, penn_emo)
penn_merged$group[penn_merged$group == 'NC'] = 'TD'
penn_merged
```

#### Make sure that the binary 'emotrics' column is all 1.

```{r}
penn_merged$bbl_id[penn_merged$emotrics == 0]
```


# Merge CHOP emotrics with demographics data

## Load CHOP emotrics data

```{r}
chop_emo = paste0('chop_final_emotrics_', latest, '.csv')
chop_emo <- read.csv(file.path(project, 
                               'Emotrics',
                               'tables',
                               latest,
                               chop_emo),
                     check.names = FALSE)

colnames(chop_emo) = cols
chop_emo = chop_emo[, grepl('bbl_id|group|Right|Left|Philtrum', 
                            colnames(chop_emo))] %>%
  drop_na()

chop_emo
```

## Merge with demographics data

```{r}
chop_merged = merge(master_demo, chop_emo)
chop_merged
```

### CHOP BBLIDs that have no demographics data

```{r}
chop_merged[is.na(chop_merged$group), ]
```
Give group values to those BBLIDs

```{r}
chop_merged$group = '22q'
```

### Make sure that emotrics values are all 1

```{r}
chop_merged[chop_merged$emotrics !=1,]
```

# Make sure that in the master demographics file 'emotrics' == 0 for bblids that don't have emotrics data

```{r}
master_demo[!master_demo$bbl_id %in% c(chop_merged$bbl_id, penn_merged$bbl_id),
            c('bbl_id', 'emotrics', 'emotrics_and_f2g')]
```

# Merge CHOP and Penn and write files

```{r}
chop_penn_sz_cr = bind_rows(penn_merged, chop_merged)

# make directory for today's date
today = format(Sys.Date(), "%m%d%y")
outdir = file.path(project, 'tables', 'Emotrics', today)
dir.create(outdir)

write.csv(chop_penn_sz_cr,
          file.path(outdir, 
                    paste0('penn_sz_cr_chop_t1_emotrics_', today, '.csv')),
          row.names = FALSE)

write.csv(chop_merged,
          file.path(outdir,
                    paste0('chop_t1_emotrics_', today, '.csv')),
          row.names = FALSE)

write.csv(penn_merged,
          file.path(outdir,
                    paste0('penn_sz_cr_emotrics_', today, '.csv')),
          row.names = FALSE)

penn_merged$group[penn_merged$group %in% c('SZ', 'CR')] = 'PS'
chop_penn = bind_rows(penn_merged, chop_merged)

write.csv(penn_merged,
          file.path(outdir,
                    paste0('penn_emotrics_', today, '.csv')),
          row.names = FALSE)

write.csv(chop_penn,
          file.path(outdir, 
                    paste0('penn_chop_t1_emotrics_', today, '.csv')),
          row.names = FALSE)

```


