---
title: "Compile F2G with Demographics"
author: "Joelle Jee"
date: "January 28, 2022"
output:
  html_document:
    df_print: paged
---

```{r}
if (!requireNamespace("tidyverse")) {
  install.packages("tidyverse")
}

if (!requireNamespace("readxl")) {
  install.packages("readxl")
}

library(tidyverse)

project = project = file.path('/Users', Sys.info()['user'], 'Library', 
                    'CloudStorage', 'Box-Box', 
                    'Craniofacial Study', 'final_emotrics_f2g_analysis')

# make directory for today's date
today = format(Sys.Date(), "%m%d%y")
outdir = file.path(project, 'tables', 'F2G', today)
dir.create(outdir)

```


# Read demographics data

```{r}

# read the most recent demographics master file
find_latest <- function(dir) {
  files = list.files(dir)
  files = gsub("F2G", "", files) %>%
    gsub("\\D", "", .) %>% 
    as.Date('%m%d%y')
  
  latest = sort(files, decreasing = TRUE)[1]
  latest = format(latest, '%m%d%y')
  
}


# read the most recent demographics master file
latest = find_latest(file.path(project, 'demographics'))
master_demo <- list.files(file.path(project, 'demographics'))
master_demo = master_demo[endsWith(master_demo, paste0(latest, '.csv'))]
master_demo <- read.csv(file.path(project, 
                                  'demographics', 
                                  master_demo), 
                        check.names = FALSE)
master_demo$`In-lab Picture` = as.numeric(master_demo$`In-lab Picture`)
master_demo
```

# Read CHOP data

## F2G data

```{r}

latest = find_latest(file.path(project, 'F2G', 'data'))
chop_fname = file.path(project, 'F2G', 'data', 
                       paste0('CHOP_F2G_demo_and_syndromes_masterfile_',
                              latest, '.xlsx'))

(chop_syndromes <- readxl::read_excel(chop_fname, sheet = "detected syndromes"))
```

## Demographics Data

## Merge the demographics and F2G data by Case ID

```{r}
chop_demo <- master_demo[grepl('22q',master_demo$group), ]
chop_demo = chop_demo[grepl('1', chop_demo$f2g),]
chop_tp1 <- merge(chop_syndromes, chop_demo, by='case_id') %>%
  select(colnames(master_demo), syndrome_name, gestalt_score)
chop_tp1
```

### Make sure that f2g columns are all 1 in the master file for those that have f2g data

```{r}
chop_tp1[chop_tp1$f2g != 1,]
```


### Make sure that f2g column in master file are 0 for those that don't have f2g data

```{r}
chop_demo[!chop_demo$case_id %in% chop_syndromes$case_id, ]
```

## BBLIDs that have less than 30 syndromes

```{r}
chop_total_count = chop_tp1 %>%
  group_by(bbl_id) %>%
  count()

chop_total_count[chop_total_count$n != 30,]
```


## Wide Version
```{r}
(chop_tp1_wide <- spread(chop_tp1, syndrome_name, gestalt_score, fill = 0))

```



```{r}

write_csv(chop_tp1,
          file.path(outdir, paste0('chop_tp1_longV_f2g_', today, '.csv')))

write.csv(chop_tp1_wide, 
          file.path(outdir, paste0('chop_tp1_wideV_f2g_', today, '.csv')),
          row.names = FALSE)


```


# Read Penn data

## Read Penn F2G data

```{r}

latest = find_latest(file.path(project, 'F2G', 'data'))
penn_fname = file.path(project, 'F2G', 'data', 
                       paste0('Penn_F2G_demo_and_syndromes_masterfile_',
                              latest, '.xlsx'))
(penn_syndrome <- readxl::read_excel(penn_fname, "detected syndromes") %>%
    rename('bbl_id' = bblid))
```

## Merge F2G and demographics data

```{r}
penn_demo = master_demo[!grepl('22q', master_demo$group),]
penn_demo = penn_demo[grepl('1', penn_demo$f2g),]
penn_total <- merge(penn_syndrome, 
                     penn_demo, 
                     by=c('bbl_id', 'case_id')) %>%
  select(colnames(master_demo), syndrome_name, gestalt_score)

penn_total$group[penn_total$group == 'NC'] = 'TD'
```

Misaligned info on demographics and syndromes data.

```{r}
penn_total_count = penn_total %>% 
  group_by(bbl_id) %>%
  count()

setdiff(penn_demo$bbl_id, penn_total_count$bbl_id)
```
Case id in demo for 10256 = 292984, in sydromes = 292985. BBLID 84332 doens't have syndromes file. F2G column in master file should be 0.


### Make sure that f2g columns are all 1 in the master file for those that have f2g data

```{r}
penn_total[penn_total$f2g != 1,]
```

### Make sure that f2g column in master file are 0 for those that don't have f2g data

```{r}
penn_demo[! penn_demo$bbl_id %in%penn_total$bbl_id, ]
```
BBLID 10256 CASE ID is 292985 in syndromes data and 292984 in demographics data. Which one is it?

### Make a version where SZ and CR are PS

```{r}

penn_total_psV = penn_total
penn_total_psV$group[penn_total_psV$group %in% c('SZ', 'CR')] = 'PS'

penn_total_psV

```


# Make wide version

Write the wide version of merged penn chart
```{r}
(penn_total_wide <- spread(penn_total, syndrome_name, gestalt_score, fill = 0))
```

```{r}
(penn_total_psV_wide <- spread(penn_total_psV, syndrome_name, gestalt_score, fill = 0))
```


## Write files

```{r}
write.csv(penn_total_wide, 
          file.path(outdir, paste0('penn_sz_cr_wideV_f2g_', today, '.csv')),
          row.names = FALSE)

write.csv(penn_total, 
          file.path(outdir, paste0('penn_sz_cr_longV_f2g_', today, '.csv')),
          row.names = FALSE)

write.csv(penn_total_psV_wide, 
          file.path(outdir, paste0('penn_wideV_f2g_', today, '.csv')),
          row.names = FALSE)

write.csv(penn_total_psV,
          file.path(outdir, paste0('penn_longV_f2g_', today, '.csv')),
          row.names = FALSE)
```

```{r}
penn_chop_tp1_psV_wide = bind_rows(penn_total_psV_wide,
                                   chop_tp1_wide)
syndromes = penn_chop_tp1_psV_wide[, !colnames(penn_chop_tp1_psV_wide) %in% colnames(master_demo)]
syndromes[is.na(syndromes)] <- 0
penn_chop_tp1_psV_wide[, !colnames(penn_chop_tp1_psV_wide) %in% colnames(master_demo)] = syndromes


write.csv(penn_chop_tp1_psV_wide,
          file.path(outdir, paste0('penn_chop_tp1_wideV_f2g_', today, '.csv')),
          row.names = FALSE)

```


```{r}

write.csv(bind_rows(penn_total_psV,
                    chop_tp1),
          file.path(outdir, paste0('penn_chop_tp1_longV_f2g_', today, '.csv')),
          row.names = FALSE)

write.csv(bind_rows(penn_total,
                    chop_tp1),
          file.path(outdir, paste0('penn_szcr_chop_tp1_longV_f2g_', today, '.csv')),
          row.names = FALSE)
```

```{r}
penn_chop_tp1_wide = bind_rows(penn_total_wide,
                                   chop_tp1_wide)
syndromes = penn_chop_tp1_wide[, !colnames(penn_chop_tp1_wide) %in% colnames(master_demo)]
syndromes[is.na(syndromes)] <- 0
penn_chop_tp1_wide[, !colnames(penn_chop_tp1_wide) %in% colnames(master_demo)] = syndromes

write.csv(bind_rows(penn_total_wide,
                    chop_tp1_wide),
          file.path(outdir, paste0('penn_szcr_chop_tp1_wideV_f2g_', today, '.csv')),
          row.names = FALSE)
```